## The code for GBM model of house price (medv) using the BostonHousing data

Load the libraries and the BostonHousing data.

```{r}
library(caret)
library(gbm)
library(tidyverse)
library(mlbench)

data(BostonHousing)
head(BostonHousing)
```
```{r}
r = data.frame(summary(BostonHousing))
write.csv(BostonHousing,"1.csv")
```

Split the data into training and testing (validation) subsets using the createDataPartition function;

```{r}
set.seed(1234) # set the seed

train.index = createDataPartition(BostonHousing$medv,p=0.7,list=F)
data.train = BostonHousing[train.index,]
data.test = BostonHousing[-train.index,]
```
The distributions of the target variable are similar across the 2 splits:

```{r}
summary(data.train$medv)
summary(data.test$medv)
```

Rescale the training and validation subsets, and keep the target variable in its original form.

```{r}
data.train.z =
data.train %>% select(-medv)  %>%
mutate_if(is_logical,as.character)  %>%
mutate_if(is_double,scale)  %>%   data.frame()
data.test.z  =
data.test  %>% select(-medv)  %>%
mutate_if(is_logical,as.character)  %>%
mutate_if(is_double,scale)  %>%   data.frame()
#  add unscaled  y variable  back
data.train.z$medv = data.train$medv
data.test.z$medv = data.test$medv
```

GBM model

```{r}
modelLookup("gbm")
```
A tuning grid can be set for these, and evaluation metric defined:
```{r}
caretGrid  <-  expand.grid(interaction.depth=c(3,5,7,10),  n.trees= (0:50)*50,
shrinkage=c(0.01 ,0.001),
n.minobsinnode=20)
metric  <-"RMSE"

trainControl<- trainControl (method="cv" , number=5)
```
Then the model can be run over the grid, setting a seed for reproducibility:

```{r}
set.seed(99)
gbm.caret  <-  train(medv ~.,data=data.train.z,  distribution="gaussian" , method="gbm" ,
trControl=trainControl, verbose=FALSE ,
tuneGrid=caretGrid, metric=metric)
```

explore  the  results
```{r}
print (gbm.caret)
ggplot(gbm.caret)
```
```{r}
names (gbm.caret)
```
```{r}
# see  best  tune
gbm.caret[6]
```

```{r}
# see  grid  results
head(data.frame(gbm.caret[4]))
dim(caretGrid)
dim(data.frame(gbm.caret[4]))
```
Find  the  best parameter  COmbinatiOn
```{r}
#put  into  a  data.frame
grid_df = data.frame(gbm.caret[4])
grid_df[which.min(grid_df$results.RMSE),]
```
```{r}
# assign  to  params  ana  inspect
params  =  grid_df[which.min(grid_df$results.RMSE),  1:4]
params
```
These can be used in the final model:
```{r}
##  Create  final  mOdel
# because  parameters  are  known,  model  can  be  fit  without  parameter  tuning
fitControl  <-  trainControl (method  = "none" ,  classProbs=FALSE)
# extract the values  from  params
gbmFit  <-  train(medv~ .,  data=data.train.z,  distribution="gaussian" , method= "gbm", trControl  =  fitControl,verbose = FALSE ,
##  Only  a  single  model  is  passed  to  the
tuneGrid  =  data.frame(interaction.depth = 10 ,
n.trees =  2500 ,
shrinkage =  0.01 ,
n.minobsinnode =  20),
metric = metric)

##  Prediction  and  model  evaluation
# generate  predictions
pred= predict(gbmFit, newdata  =data.test.z)
# plot  these  against  observed
data.frame(Predicted  =  pred,  Observed  =  data.test.z$medv)  %>%
ggplot(aes(x =Observed, y = Predicted))+  geom_point (size =1,  alpha = 0.5)+geom_smooth(method ="lm")
```

```{r}
# generate  some  prediction  accuracy  measures
postResample (pred= pred,  obs=data.test.z$medv)

# examine  variable  importance
varImp(gbmFit, scale  =  FALSE)

```
compare with the OLS:
```{r}
##  Create  final  mOdel
# because  parameters  are  known,  model  can  be  fit  without  parameter  tuning
fitControl  <-  trainControl (method  = "none" ,  classProbs=FALSE)
# extract the values  from  params
OLSFit  <-  train(medv~ .,  data=data.train.z,  distribution="gaussian" , method= "lm", trControl  =  fitControl,verbose = FALSE ,
metric = metric)

##  Prediction  and  model  evaluation
# generate  predictions
pred_OLS= predict(OLSFit, newdata  =data.test.z)
postResample (pred= pred_OLS,  obs=data.test.z$medv)
```

```{r}
# plot  these  against  observed
data.frame(Predicted  =  pred_OLS,  Observed  =  data.test.z$medv)  %>%
ggplot(aes(x =Observed, y = Predicted))+  geom_point (size =1,  alpha = 0.5)+geom_smooth(method ="lm")
```

